{\large Gravitational wave sources in our Galactic backyard:}\\{\normalsize Predictions for BHBH, BHNS and NSNS binaries in LISA

### 2021-10-16 paper/tex/sections/appendix.tex
further fixes to figure refs/links

\begin{figure}[htb]
    \centering
    \includegraphics[width=\columnwidth]{fig15_mBH_simple_mw_variation.pdf}
    \caption{As Fig.~\ref{fig:fiducial_pdf_distributions} (top left panel), but for the Milky Way model used in \citet{Breivik+2020}. Dotted lines show the distribution from Fig.~\ref{fig:fiducial_pdf_distributions} for comparison. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig15_mBH_simple_mw_variation.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/fiducial.ipynb}{\faBook}.}
    \label{fig:bh_mass_simple_mw}
\end{figure}

@@

% \subsection{Distribution of eccentric sources on sensitivity curve}
\begin{figure*}[b]
    \centering
    \includegraphics[width=\textwidth]{fig17_dcos_on_sc_eccentric_colours.png}
    \caption{As the bottom panels of Fig.~\ref{fig:dcos_on_sc}, but without the density distributions and scatter points are coloured by their eccentricity. We show eccentric sources are located in an offshoot below the $30 \unit{kpc}$ around $2 \unit{mHz}$. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig17_dcos_on_sc_eccentric_colours.png}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/sensitivity_curve.ipynb}{\faBook}.}
    \label{fig:dcos_on_sc_ecc_col}
\end{figure*}

% \subsection{Properties of detectable systems (progenitors and at DCO formation)}
\begin{figure*}[ht]
    \centering
    \includegraphics[width=0.66\textwidth]{fig18_dco_formation_distributions.pdf}
    \caption{As Fig.~\ref{fig:fiducial_pdf_distributions}, but for the properties of the progentiors of detectable systems (top row) and the properties of the detectable systems at DCO formation (bottom row). Note that the $m_{1, {\rm ZAMS}}$ panel uses a log scaled $y$-axis. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig18_dco_formation_distributions.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/fiducial.ipynb}{\faBook}.}
    \label{fig:progenitor_and_dco_properties}
\end{figure*}

% \subsection{Detection rates relative to fiducial model}
\begin{figure*}[hb]
    \centering
    \includegraphics[width=\textwidth]{fig19_dco_relative_rates.pdf}
    \caption{Similar to Fig.~\ref{fig:detection_rates}, this shows the \textit{relative} detection rates compared to the fiducial model in order to emphasise how much each DCO is affected by model variations compared to other DCOs. The scatter points show the mean and the shaded areas show the 1-$\sigma$ uncertainty range. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig19_dco_relative_rates.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/detections.ipynb}{\faBook}.}
    \label{fig:dco_relative_rates}
\end{figure*}

>>>>

\begin{figure}[htb]
    \centering
    \includegraphics[width=\columnwidth]{fig15_mBH_simple_mw_variation.pdf}
    \caption{As Fig.~\ref{fig:fiducial_pdf_distributions}b, but for the Milky Way model used in \citet{Breivik+2020}. Dotted lines show the distribution from Fig.~\ref{fig:fiducial_pdf_distributions}b for comparison. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig15_mBH_simple_mw_variation.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/fiducial.ipynb}{\faBook}.}
    \label{fig:bh_mass_simple_mw}
\end{figure}

@@

% \subsection{Distribution of eccentric sources on sensitivity curve}
\begin{figure*}[p]
    \centering
    \includegraphics[width=\textwidth]{fig17_dcos_on_sc_eccentric_colours.png}
    \caption{As the bottom panels of Fig.~\ref{fig:dcos_on_sc}, but without the density distributions and scatter points are coloured by their eccentricity. We show eccentric sources are located in an offshoot below the $30 \unit{kpc}$ around $2 \unit{mHz}$. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig17_dcos_on_sc_eccentric_colours.png}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/sensitivity_curve.ipynb}{\faBook}.}
    \label{fig:dcos_on_sc_ecc_col}

    \includegraphics[width=0.8\textwidth]{fig18_dco_formation_distributions.pdf}
    \caption{As Fig.~\ref{fig:fiducial_pdf_distributions}, but for the properties of the detectable systems at DCO formation. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig18_dco_formation_distributions.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/fiducial.ipynb}{\faBook}.}
    \label{fig:dco_formation_properties}

    \includegraphics[width=\textwidth]{fig19_dco_relative_rates.pdf}
    \caption{Similar to Fig.~\ref{fig:detection_rates}, this shows the \textit{relative} detection rates compared to the fiducial model in order to emphasise how much each DCO is affected by model variations compared to other DCOs. The scatter points show the mean and the shaded areas show the 1-$\sigma$ uncertainty range. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig19_dco_relative_rates.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/detections.ipynb}{\faBook}.}
    \label{fig:dco_relative_rates}
\end{figure*}

### 2021-10-16 paper/tex/sections/comparisons.tex
further fixes to figure refs/links

    \includegraphics[width=\textwidth]{fig13_2_compare_dco.png}
    \caption{A table comparing previous studies of a similar nature to this work. The works listed in the table are \citet{Nelemans+2001}, \citet{Belczynski+2010}, \citet{Liu+2014}, \citet{Lamberts+2018}, \citet{Sesana+2020}, \citet{Lau+2020}, \citet{Breivik+2020} and \citet{Shao+2021}. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig13_1_compare_dco.png}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig13_w_compare_dco.png}{\faFileImage}.}
    \label{fig:previous_studies}
\end{figure*}

>>>>

    \includegraphics[width=\textwidth]{fig13_2_compare_dco.png}
    \caption{A table comparing previous studies of a similar nature to this work. The works listed in the table are \citet{Nelemans+2001}, \citet{Belczynski+2010}, \citet{Liu+2014}, \citet{Lamberts+2018}, \citet{Sesana+2020}, \citet{Lau+2020}, \citet{Breivik+2020} and \citet{Shao+2021}. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig13_1_compare_dco.png}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig13_2_compare_dco.png}{\faFileImage}.}
    \label{fig:previous_studies}
\end{figure*}

### 2021-10-16 paper/tex/sections/results.tex
further fixes to figure refs/links

A further reason for the preference for low masses is that the population of BHs detectable by LISA primarily come from progenitors that formed from high metallicity gas according to our simulations (see Sec.~\ref{sec:progenitors_and_formation} also Fig.~\ref{fig:progenitor_and_dco_properties}). Stellar winds are stronger at high metallicity leading to increased mass loss. This affects the mass of the most massive black holes that can be formed \citep{Belczynski+2010}.

@@

When comparing the distribution of formation times for the three different DCO types we see that NSNSs are most strongly concentrated at recent times, followed by BHBH and then BHNS. To understand this it is helpful to consider how the inspiral time scales with various parameters
\begin{equation}
    t_{\rm inspiral} \propto \frac{a^4}{(m_1 + m_2)^{3}} \cdot \frac{q}{(1+q)^2} \cdot (1 - e^2)^{-7/2}.
\end{equation}
The inspiral time depends most strongly on the separation at DCO formation ($a$) and this is where the three types also differ most strongly (see Fig.~\ref{fig:progenitor_and_dco_properties}). The NSNS systems have the tightest orbits at DCO formation. The median of the distribution of separations at DCO formation, $\langle a_{\rm DCO} \rangle_{\rm med}$, relate as 8:3:1 for BHBH:BHNS:NSNS in our simulations. This results in increase of the inspiral time by a factor of about 4000:80:1. The total masses affect the inspiral time to the third power and this where the heavier BHBH systems are favoured. The median total masses differ by ratios of 6:4:1 for BHBH:BHNS:NSNS in our simulations, impacting the inspiral times such that they are a factor of 200:60:1 shorter, partially counteracting the effect of the separations. The term depending on the mass ratio $q$ only varies by about 30\% for the mass ratio ranges considered here and so is not of interest. The eccentricity term is not of importance for mildly eccentric systems, $f(e_{\rm DCO} \le 0.3)  \le 1.4$ but of large importance for the very eccentric  $f(e_{\rm DCO} \ge 0.9) \ge 300$. The fraction of highly eccentric systems with $e_{\rm DCO}>0.9$ is 33\%, 16\$ and 8\% of for BHBH, BHNS and NSNS respectively, see also Fig.~\ref{fig:progenitor_and_dco_properties}.

@@

\paragraph{Time until merger} The final panel of Fig.~\ref{fig:fiducial_pdf_distributions} shows the remaining time until merger for each of the DCO types at the start of LISA mission. The distributions are strikingly similar and peak with merger times of around a Myr.

@@

\subsection{Distribution in the Milky Way}\label{sec:mw_detectable_distribution}
For our Milky way model we considered three different components, a low-$\alpha$ (``thin'') disc, an older high-$\alpha$ (``thick'') disc and a bulge/bar. In Table~\ref{tab:component_fractions} we summarise the number of detections originating from each of these components. Despite the fact that only $42.5\%$ of systems are formed in the low-$\alpha$ disc, we find that $86\%$ of the detections originate from this component. This is because most detectable systems were formed relatively recently (see Fig.~\ref{fig:fiducial_pdf_distributions}) and so the high-$\alpha$ disc and bulge are effectively too old to contribute many detectable systems. Nevertheless, we do find a significant fraction of detectable systems originate in the high-$\alpha$ disc and bulge, indicating that it is still important to include these components, as was ignored in some earlier works (see Sec.~\ref{sec:compare_studies}).

@@

\begin{table}[tb]
    \centering
    \begin{tabular}{l|ccccc}
        \hline
        \multirow{2}{*}{Component} & Formation & \multicolumn{4}{c}{Detectable} \\ \cline{2-6}
        & {\scriptsize All} & {\scriptsize All} & {\scriptsize BHBH} & {\scriptsize BHNS} & {\scriptsize NSNS} \\
        \hline
        Low-$\alpha$ disc & 42.5\% & 86\% & 89\% & 82\% & 85\%\\
        High-$\alpha$ disc & 42.5\% & 6\% & 5\% & 8\% & 12\%\\
        Bulge/bar & 15.0\% & 8\% & 6\% & 10\% & 3\%\\
    \end{tabular}
    \caption{Percentage of systems in each Galactic component.}
    \label{tab:component_fractions}
\end{table}
\begin{figure}[tb]
    \centering
    \includegraphics[width=\columnwidth]{fig4_1_galaxy_density_distribution.png}
    \includegraphics[width=\columnwidth]{fig4_2_detectable_distance_distribution_4yr.pdf}
    \caption{
    \textbf{Top:} A face-on view of the Galactic density distribution for detectable DCOs. We show the density distribution for the top 95\% of the sources, the rest are indicated by scatter points whose sizes correspond to their sampling weights. The coloured lines show the 75\% contour for each of the individual DCO types. The large cross passes through (0, 0) and helps to highlight the bias towards the position of the sun, which is indicated by the $\odot$.
    \textbf{Bottom:}   As Figure~\ref{fig:fiducial_pdf_distributions}, but for the luminosity distance. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig4_2_galaxy_density_distribution.png}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/galaxy_creation_station.ipynb}{\faBook} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig4_1_detectable_distance_distribution_4yr.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/fiducial.ipynb}{\faBook}.}
    \label{fig:detectable_distance_dist}
\end{figure}

@@

\begin{figure}[htb]
    \centering
    \includegraphics[width=\columnwidth]{fig5_progenitor_distributions.pdf}
    \caption{As Fig.~\ref{fig:fiducial_pdf_distributions}, but for the initial primary mass (top) and inital mass ratio (bottom). \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig5_progenitor_distributions.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/fiducial.ipynb}{\faBook}.}
    \label{fig:progenitor_properties}
\end{figure}

>>>>

A further reason for the preference for low masses is that the population of BHs detectable by LISA primarily come from progenitors that formed from high metallicity gas according to our simulations. Stellar winds are stronger at high metallicity leading to increased mass loss. This affects the mass of the most massive black holes that can be formed \citep{Belczynski+2010}.

@@

When comparing the distribution of formation times for the three different DCO types we see that NSNSs are most strongly concentrated at recent times, followed by BHBH and then BHNS. To understand this it is helpful to consider how the inspiral time scales with various parameters
\begin{equation}
    t_{\rm inspiral} \propto \frac{a^4}{(m_1 + m_2)^{3}} \cdot \frac{q}{(1+q)^2} \cdot (1 - e^2)^{-7/2}.
\end{equation}
The inspiral time depends most strongly on the separation at DCO formation ($a$) and this is where the three types also differ most strongly (see Fig.~\ref{fig:dco_formation_properties}). The NSNS systems have the tightest orbits at DCO formation. The median of the distribution of separations at DCO formation, $\langle a_{\rm DCO} \rangle_{\rm med}$, relate as 8:3:1 for BHBH:BHNS:NSNS in our simulations. This results in increase of the inspiral time by a factor of about 4000:80:1. The total masses affect the inspiral time to the third power and this where the heavier BHBH systems are favoured. The median total masses differ by ratios of 6:4:1 for BHBH:BHNS:NSNS in our simulations, impacting the inspiral times such that they are a factor of 200:60:1 shorter, partially counteracting the effect of the separations. The term depending on the mass ratio $q$ only varies by about 30\% for the mass ratio ranges considered here and so is not of interest. The eccentricity term is not of importance for mildly eccentric systems, $f(e_{\rm DCO} \le 0.3)  \le 1.4$ but of large importance for the very eccentric  $f(e_{\rm DCO} \ge 0.9) \ge 300$. The fraction of highly eccentric systems with $e_{\rm DCO}>0.9$ is 33\%, 16\$ and 8\% of for BHBH, BHNS and NSNS respectively, see also Fig.~\ref{fig:dco_formation_properties}.

@@

\paragraph{Time until merger} Fig.~\ref{fig:fiducial_pdf_distributions}f shows the remaining time until merger for each of the DCO types at the start of LISA mission. The distributions are strikingly similar and peak with merger times of around a Myr.

@@

\subsection{Distribution in the Milky Way}\label{sec:mw_detectable_distribution}
For our Milky way model we considered three different components, a low-$\alpha$ (``thin'') disc, an older high-$\alpha$ (``thick'') disc and a bulge/bar. In Table~\ref{tab:component_fractions} we summarise the number of detections originating from each of these components. Despite the fact that only $42.5\%$ of systems are formed in the low-$\alpha$ disc, we find that $86\%$ of the detections originate from this component. This is because most detectable systems were formed relatively recently (see Fig.~\ref{fig:fiducial_pdf_distributions}e) and so the high-$\alpha$ disc and bulge are effectively too old to contribute many detectable systems. Nevertheless, we do find a significant fraction of detectable systems originate in the high-$\alpha$ disc and bulge, indicating that it is still important to include these components, as was ignored in some earlier works (see Sec.~\ref{sec:compare_studies}).

@@

\begin{table}[tb]
    \centering
    \begin{tabular}{l|ccccc}
        \hline
        \multirow{2}{*}{Component} & Formation & \multicolumn{4}{c}{Detectable} \\ \cline{2-6}
        & {\scriptsize All} & {\scriptsize All} & {\scriptsize BHBH} & {\scriptsize BHNS} & {\scriptsize NSNS} \\
        \hline
        Low-$\alpha$ disc & 42.5\% & 86\% & 89\% & 82\% & 85\%\\
        High-$\alpha$ disc & 42.5\% & 6\% & 5\% & 8\% & 12\%\\
        Bulge/bar & 15.0\% & 8\% & 6\% & 10\% & 3\%\\
    \end{tabular}
    \caption{Percentage of systems in each Galactic component.}
    \label{tab:component_fractions}
\end{table}
\begin{figure}[tb]
    \centering
    \includegraphics[width=\columnwidth]{fig4_1_galaxy_density_distribution.png}
    \includegraphics[width=\columnwidth]{fig4_2_detectable_distance_distribution_4yr.pdf}
    \caption{
    \textbf{Top:} A face-on view of the Galactic density distribution for detectable DCOs. We show the density distribution for the top 95\% of the sources, the rest are indicated by scatter points whose sizes correspond to their sampling weights. The coloured lines show the 75\% contour for each of the individual DCO types. The large cross passes through (0, 0) and helps to highlight the bias towards the position of the sun, which is indicated by the $\odot$.
    \textbf{Bottom:} As Figure~\ref{fig:fiducial_pdf_distributions}, but for the luminosity distance. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig4_1_galaxy_density_distribution.png}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/galaxy_creation_station.ipynb}{\faBook} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig4_2_detectable_distance_distribution_4yr.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/fiducial.ipynb}{\faBook}.}
    \label{fig:detectable_distance_dist}
\end{figure}

@@

\begin{figure}[htb]
    \centering
    \includegraphics[width=\columnwidth]{fig5_progenitor_distributions.pdf}
    \caption{As Fig.~\ref{fig:fiducial_pdf_distributions}, but for the initial primary mass (top) and inital mass ratio (bottom). Note that the y-axis is log-scaled for the initial primary mass. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig5_progenitor_distributions.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/fiducial.ipynb}{\faBook}.}
    \label{fig:progenitor_properties}
\end{figure}

### 2021-10-16 paper/tex/sections/variations.tex
further fixes to figure refs/links

\begin{figure*}[p]
    \centering
    \includegraphics[width=\textwidth]{fig9_dco_detections.pdf}
    \caption{The number of expected detections in the LISA mission for different DCO types and model variations. Error bars show the 1- (solid) and 2-$\sigma$ (dotted) Poisson uncertainties. An arrow indicates that the error bar extends to zero. The left axis and grid lines show the number of detections in a 4-year LISA mission and the right axis shows an approximation of the number of detections in a 10-year mission (we scale the axis by $\sqrt{T_{\rm obs}}$, see Table~\ref{tab:detection_rates} for exact rates). Each model is described in further detail in Table~\ref{tab:physics_variations} and details of the fiducial assumptions are in Section~\ref{app:fiducial_physics}. See also Fig.~\ref{fig:dco_relative_rates} and Sec.~\ref{sec:detection_rate_analysis} for a discussion. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig9_dco_detections.png}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/detections.ipynb}{\faBook}.}
    \label{fig:detection_rates}
\end{figure*}

>>>>

\begin{figure*}[p]
    \centering
    \includegraphics[width=\textwidth]{fig9_dco_detections.pdf}
    \caption{The number of expected detections in the LISA mission for different DCO types and model variations. Error bars show the 1- (solid) and 2-$\sigma$ (dotted) Poisson uncertainties. An arrow indicates that the error bar extends to zero. The left axis and grid lines show the number of detections in a 4-year LISA mission and the right axis shows an approximation of the number of detections in a 10-year mission (we scale the axis by $\sqrt{T_{\rm obs}}$, see Table~\ref{tab:detection_rates} for exact rates). Each model is described in further detail in Table~\ref{tab:physics_variations} and details of the fiducial assumptions are in Section~\ref{app:fiducial_physics}. See also Fig.~\ref{fig:dco_relative_rates} and Sec.~\ref{sec:detection_rate_analysis} for a discussion. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig9_dco_detections.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/detections.ipynb}{\faBook}.}
    \label{fig:detection_rates}
\end{figure*}
